Linear ADRC is equivalent to PID with set-point weighting and measurement filter

### 2025-01-18 paper/main.tex
include statespace realization of equivalent controllers

\documentclass[letterpaper, 10 pt, conference]{ieeeconf} 
\IEEEoverridecommandlockouts                              % This command is only needed if 
                                                          % you want to use the \thanks command
\overrideIEEEmargins                                      % Needed to meet printer requirements.
\usepackage{amsmath}

\usepackage{graphicx} % Required for inserting images
\usepackage{hyperref}
\hypersetup{
	colorlinks=true,
	linkcolor=black,
	citecolor=black,
	filecolor=black,
	urlcolor=black,
}
\usepackage{cleveref}
\crefformat{footnote}{#2\footnotemark[#1]#3}

@@

\section{Introduction}
Active Disturbance-Rejection Control (ADRC) is often touted as a promising improvement over the PID controller as the goto tool for practitioners. In \cite{herbst2013simulative}, the author point out that ADRC, in particular linear ADRC, is a new name for the well-established method of using state feedback with state estimates provided by an observer augmented with a disturbance model. Linear ADRC makes the particular case of an integrating disturbance model acting on the plant input.

@@

In this paper, we show that linear ADRC tuned with the bandwidth method is in fact equivalent to a commonplace implementation of a PID controller. For \emph{first-order systems}, the controller response to measurements is equivalent to a PI controller with a first-order filter on the measurement signal, and the response to references is a PID controller(!). With very minor approximation error, the complete two degree-of-freedom controller\footnote{Two inputs $r$ and $y$ with different response to each input.} can be represented as a single PI controller with set-point weighting and a first-order lowpass filter on the measurement, a very convenient conclusion considering how commonplace this particular implementation of a PI controller is. For \emph{second-order systems}, with a similar minor approximation (in reference response only, no approximation is made in the response from measurements), the linear ADRC controller is equivalent to a PID controller with set-point weighting and a second-order lowpass filter on measurements. Once again, a very convenient conclusion considering the ubiquity of this particular implementation.

@@

\section{details}
We will follow the notation used in \cite{herbst2013simulative}, that considers an ADRC controller on the following form
\begin{align}
   \bmatrixx{\dot{\hat{x}}_1 \\ \dot{\hat{x}}_2} &= 
   		\bmatrixx{-l_1 & 1 \\ -l_2 & 0} \bmatrixx{\hat{x}_1 \\ \hat{x}_2} + 
		\bmatrixx{b_0 \\ 0} u + \bmatrixx{l_1 \\ l_2} y \\
	u &= \dfrac{-K_P(r-y) - \hat{f}}{b_0} = \dfrac{-K_P(r-\hat{x}_1) - \hat{x}_2}{b_0}
\end{align}
where $r$ is the reference signal, $y$ is the measurement signal, $u$ is the control signal, and $\hat{f}$ is the estimated disturbance input. The tuning parameters are $l_1$, $l_2$, and $K_P$, and these are set by picking a desired settling time $T_s$ and the \emph{characteristic gain} of the plant, $b_0$. The plant is assumed to be on the form
$$P(s) = \dfrac{K}{Ts+1}$$
which gives $b_0 = K/T$. The parameters as dictated by the tuning rule are selected as 
% Pdes1 = ss(tf([1],[1, 0]))
% Pdes2 = add_low_frequency_disturbance(Pdes1, 1)

@@

\subsection{First-order linear ADRC as a transfer function}
When represented as a transfer function, the ADRC controller detailed above can be written as\footnote{We omit the parameter $b_0$ here for brevity. It scales the controller gain proportionally and is thus straightforward to incorporate.}

@@

\begin{align}
	C_{ADRC}(s) = \bmatrixx{C_r(s) & C_y(s)}\\
	C_r(s) = \dfrac{4T_{s}^2s^2 + 32T_{s}gs + 64g^2}{T_{s}^3s^2 + T_{s}^2(8g + 4)s} \\
	C_y(s) = \dfrac{- 32T_{s}gs -(16T_{s} + 64)g^2}{s(T_{s}^3s + T_{s}^2(8g + 4))}
\end{align}
(note how $C_y$ has negative feedback built in). Compare $C_y(s)$ to the transfer function of a PI controller with lowpass filter
$$C_{PIF} = \big(k_p + k_i/s \big)\dfrac{1}{T_f s + 1} = \dfrac{k_p s + k_i}{s(T_f s + 1)}$$
it becomes clear that we can pick $k_p, k_i, T_f$ so that the PI controller becomes identical to the feedback part of the ADRC controller, $C_y$:
\begin{align}
	k_p &= \dfrac{-4 g^2 - 8 g}{T_s(2 g + 1)} \label{eq:kp}\\
	k_i &= \dfrac{-16 g^2}{T_s^2(2g + 1)} \label{eq:ki}\\
	T_f &= \dfrac{T_s}{8 g + 4} \label{eq:Tf}
\end{align}

The feedforward part, $C_r$, can similarly be shown to be equivalent to a first-order filtered PID controller, but more interestingly, we can approximate $C_r$ as $C_y$ with set-point weighting on the proportional term, but without the lowpass filter. To do this, we introduce the 2DOF PI controller $K_{ry}$ with set-point weight $b$, defined by the transfer function between $r,y$ and $u$ in the expression
$$K_{ry}(s) : u = k_p (br - y) + k_i(r-y)\frac{1}{s}$$
to $b = K_P / k_p$ where $K_P$ is the proportional gain in the ADRC controller and $k_p$ is the proportional gain in the equivalent PID controller obtained by matching $C_y$ to $C_{PIF}$. With this choice of $b$, the feedforward part of the 2DOF PI controller will have the following low and high-frequency asymptotes:
\begin{align}
	s &\to 0:  & \quad C_r(s) &\to \dfrac{16g^2}{T_{s}^2(2g + 1)s} \\
	s &\to 0:  & \quad K_{ry}(s) &\to \dfrac{k_i}{s} 
\end{align}

@@

\subsection{Second-order linear ADRC}
In a similar fashion we can establish equivalence between the second-order linear ADRC controller and a PID controller with set-point weighting and a second-order lowpass filter on the measurement signal, given by the transfer function between $r,y$ and $u$ in the expression:
$$K_{PIDF}(s) : u = \dfrac{k_p (br - y) + (r-y)\frac{k_i}{s} + (0r - y)k_d s}{T_f^2s^2 + 2T_f d s + 1}$$
where $d$ is a relative damping parameter of the filter and $T_f$ is the time constant of the filter. We explicitly included the zero term $0r$ in the derivative part to highlight that set-point weight in the derivative term is zero, while $b$ generally is not.

@@

\section{Numerical results} \label{sec:numerical}
This section will demonstrate the equivalence between linear ADRC and the proposed equivalent PI(D) controllers numerically, using the same plant model as the one used in \cite{herbst2013simulative}. The plant model is a first-order system with $T = 1$ and $K = 1$, and the ADRC controller is tuned with $T_s = 1$ and $g = 10$. The parameters for the PI(D) controllers are found using \cref{eq:kp,eq:ki,eq:Tf}.

@@

\subsection{First-order system}

@@

\subsection{Second-order system}
The second-order system used in the experiment section in \cite{herbst2013simulative} is given by
% K = 1
% D = 1
% T = 1
% P = tf([K],[T^2, 2D*T, 1])
$$P(s) = \dfrac{1}{T^2s^2 + 2DTs + 1} = \dfrac{1}{s^2 + 2s + 1}$$

We show the same figures as for the first-order system
\begin{itemize}
	\item \cref{fig:second_order_K} shows the step response as the value of $K$ varies.
	\item \cref{fig:second_order_T} shows the step response as the value of $T$ varies.
	\item \cref{fig:second_order_bode_C} shows the Bode plot of the controllers.
	\item \cref{fig:second_order_bode_ry} shows the Bode plot of the closed-loop response from references to output.
	\item \cref{fig:second_order_bode_uy} shows the Bode plot of the closed-loop response from measurements to control signal.
\end{itemize}
The conclusions are largely the same for second-order systems. The response from measurements is identical between the ADRC controller and the equivalent PID controller. The response to references is slightly different, visible in the bode plot, but not different enough to be clearly visible in the step responses with varying plant parameters.

@@

\section{Second-order linear ADRC as a transfer function}
The transfer functions for the second-order ADRC controller (derived with symbolic computer calculations, reproduced in the software repository \cite{repo}) can be realized by the following PID+filter parameters
\begin{align}
	k_p &= \dfrac{(-72g^3 - 108g^2)}{(3T_s^2 g^2 + 6T_s^2 g + T_s^2)} \\
	k_i &= \dfrac{-216g^3}{(3T_s^3 g^2 + 6T_s^3 g + T_s^3)} \\
	k_d &= \dfrac{(-6g^3 - 36g^2 - 18g)}{(3T_s g^2 + 6T_s g + T_s)} \\
	T_f &= T_s/6 \sqrt{1/(3g^2 + 6g + 1)} \\
	d &= (3g + 2)/2 \sqrt{1/(3g^2 + 6g + 1)} \\
	b &= -(6/T_s)^2 / k_p
\end{align}

>>>>

\documentclass[letterpaper, 10 pt, conference]{ieeeconf} 
\IEEEoverridecommandlockouts                              % This command is only needed if 
                                                          % you want to use the \thanks command
\overrideIEEEmargins                                      % Needed to meet printer requirements.
\usepackage{amsmath}
\usepackage{flushend}
\usepackage{graphicx} % Required for inserting images
\usepackage{hyperref}
\hypersetup{
	colorlinks=true,
	linkcolor=black,
	citecolor=black,
	filecolor=black,
	urlcolor=black,
}
\usepackage{cleveref}
\crefformat{footnote}{#2\footnotemark[#1]#3}

@@

\section{Introduction}
Active Disturbance-Rejection Control (ADRC) is often touted as a promising improvement over the PID controller as the goto tool for practitioners. In \cite{herbst2013simulative}, the author point out that ADRC, in particular linear ADRC, is a new name for the well-established method of using state feedback with state estimates provided by an observer augmented with a disturbance model. Linear ADRC makes the particular choice of an integrating disturbance model acting on the plant input.

@@

In this paper, we show that linear ADRC tuned with the bandwidth method is in fact equivalent to a commonplace implementation of a PID controller. For \emph{first-order systems}, the controller response to measurements is equivalent to a PI controller with a first-order filter on the measurement signal, and the response to references is a filtered PID controller. With very minor approximation error, the complete two degree-of-freedom\footnote{Two inputs $r$ and $y$ with different response to each input.} controller can be represented as a single PI controller with set-point weighting and a first-order lowpass filter on the measurement, a very convenient conclusion considering how commonplace this particular implementation of a PI controller is. For \emph{second-order systems}, with a similar minor approximation (in reference response only, no approximation is made in the response from measurements), the linear ADRC controller is equivalent to a PID controller with set-point weighting and a second-order lowpass filter on measurements. Once again, a very convenient conclusion considering the ubiquity of this particular implementation.

@@

\section{Technical Details}
We will follow the notation used in \cite{herbst2013simulative} for first-order systems, that considers an ADRC controller on the following form
\begin{align}
   \bmatrixx{\dot{\hat{x}}_1 \\ \dot{\hat{x}}_2} &= 
   		\bmatrixx{-l_1 & 1 \\ -l_2 & 0} \bmatrixx{\hat{x}_1 \\ \hat{x}_2} + 
		\bmatrixx{b_0 \\ 0} u + \bmatrixx{l_1 \\ l_2} y \\
	u &= \dfrac{-K_P(r-y) - \hat{f}}{b_0} = \dfrac{-K_P(r-\hat{x}_1) - \hat{x}_2}{b_0}
\end{align}
where the state-space system represents the disturbance observer, $r$ is the reference signal, $y$ is the measurement signal, $u$ is the control signal, and $\hat{f}$ is the estimated disturbance input. The tuning parameters are the observer parameters $l_1$, $l_2$, and $K_P$, and these are set using rules that require the user to pick a desired settling time $T_s$ and the \emph{characteristic gain} of the plant, $b_0$. The plant is assumed to be on the form
$$P(s) = \dfrac{K}{Ts+1}$$
which gives $b_0 = K/T$. The parameters as dictated by the bandwidth tuning rule are selected as 
% Pdes1 = ss(tf([1],[1, 0]))
% Pdes2 = add_low_frequency_disturbance(Pdes1, 1)

@@

\subsection{First-order linear ADRC and PID}
\subsubsection{Transfer functions}

When represented as a transfer function, the ADRC controller detailed above can be written as\footnote{We omit the parameter $b_0$ here for brevity. It scales the controller gain proportionally and is thus straightforward to incorporate.}

@@

\begin{align}
	C_{ADRC}(s) = \bmatrixx{C_r(s) & -C_y(s)}\\
	C_r(s) = \dfrac{4T_{s}^2s^2 + 32T_{s}gs + 64g^2}{T_{s}^3s^2 + T_{s}^2(8g + 4)s} \\
	C_y(s) = \dfrac{32T_{s}gs + (16T_{s} + 64)g^2}{s(T_{s}^3s + T_{s}^2(8g + 4))}
\end{align}
 Compare $C_y(s)$ to the transfer function of a PI controller with lowpass filter
$$C_{PIF} = \big(k_p + k_i/s \big)\dfrac{1}{T_f s + 1} = \dfrac{k_p s + k_i}{s(T_f s + 1)}$$
it becomes clear that we can pick $k_p, k_i, T_f$ so that the PI controller becomes identical to the feedback part of the ADRC controller, $C_y$:
\begin{align}
	k_p &= \dfrac{4 g^2 + 8 g}{T_s(2 g + 1)} \label{eq:kp}\\
	k_i &= \dfrac{16 g^2}{T_s^2(2g + 1)} \label{eq:ki}\\
	T_f &= \dfrac{T_s}{8 g + 4} \label{eq:Tf}
\end{align}

The feedforward part, $C_r$, can similarly be shown to be equivalent to a first-order filtered PID controller, but more interestingly, we can approximate $C_r$ as $-C_y$ with set-point weighting on the proportional term, but without the lowpass filter. To do this, we introduce the 2DOF PI controller $K_{ry}$ with set-point weight $b$, defined by the transfer function between $r,y$ and $u$ in the expression
$$K_{ry}(s) : u = k_p (br - y) + k_i(r-y)\frac{1}{s}$$
to $b = K_P / k_p$ where $K_P$ is the proportional gain in the ADRC controller and $k_p$ is the proportional gain in the equivalent PID controller obtained by matching $C_y$ to $C_{PIF}$. With this choice of $b$, the feedforward part of the 2DOF PI controller will have the following low and high-frequency asymptotes:
\begin{align}
	s &\to 0:  & \quad C_r(s) &\to \dfrac{16g^2}{T_{s}^2(2g + 1)s} = \dfrac{k_i}{s}  \\
	s &\to 0:  & \quad K_{ry}(s) &\to \dfrac{k_i}{s} 
\end{align}

@@

\subsubsection{State-space realization of equivalent PIF controller}
For convenience, we also include a statespace realization of the proposed equivalent 2DOF PIF controller (with the approximation of $C_r$):
\begin{align*}
		\dot{x} &= \left[\begin{matrix}0.0 & - \frac{64 g^{2}}{T_{s}^{3}}\\0 & \frac{4 \left(- 2 g - 1\right)}{T_{s}}\end{matrix}\right] x + 		
		\left[\begin{matrix}\frac{16 g^{2}}{T_{s}^{2} \left(2 g + 1\right)} & 0\\0 & 1\end{matrix}\right] \bmatrixx{r \\ y} \\
		u &= \left[\begin{matrix}1.0 & \frac{16 g \left(- g - 2\right)}{T_{s}^{2}}\end{matrix}\right] x + 		
		\left[\begin{matrix}\frac{4}{T_{s}} & 0\end{matrix}\right] \bmatrixx{r \\ y}
\end{align*}
with inputs $[r, y]$ and output $u$.

\subsection{Second-Order Linear ADRC}
In a similar fashion we can establish equivalence between the second-order linear ADRC controller and a PID controller with set-point weighting and a second-order lowpass filter on the measurement signal, given by the transfer function between $r,y$ and $u$ in the expression:
\begin{align}
	K_{PIDF}(s) &: u = k_p (br - y_f) + (r-y_f)\frac{k_i}{s} - y_fk_d s \\
	y_f &= \dfrac{1}{T_f^2s^2 + 2T_f d s + 1} y	
\end{align}
where $d$ is a relative damping parameter of the filter and $T_f$ is the time constant of the filter. Note that the reference $r$ does not appear in the derivative part, the set-point weight in the derivative term is zero, while $b$ generally is not.

@@

\subsubsection{State-space realization of equivalent PIDF controller}
The equivalent 2DOF filtered PID controller for second-order systems, including the approximation of $C_r$, can be realized in state-space form as
\begin{align*}
	\dot{x} &= \left[\begin{matrix}0 & 0 & 1\\1 & 0 & 0\\\frac{- 108 g^{2} - 216 g - 36}{T_{s}^{2}} & 0 & \frac{- 18.0 g - 12.0}{T_{s}}\end{matrix}\right]x + \\
	~ &~ \left[\begin{matrix}0 & 0\\1 & 0\\0 & \frac{- 108 g^{2} - 216 g - 36}{T_{s}^{2}}\end{matrix}\right] \bmatrixx{r \\ y} \\
	u &= \\
	~ &~ \left[\begin{matrix}\frac{g^{2} \left(72 g + 108\right)}{T_{s}^{2} \left(3 g^{2} + 6 g + 1\right)} & \frac{216 g^{3}}{T_{s}^{3} \left(3 g^{2} + 6 g + 1\right)} & \frac{g \left(6 g^{2} + 36 g + 18\right)}{T_{s} \left(3 g^{2} + 6 g + 1\right)}\end{matrix}\right]x \\ 
	~ &+ \left[\begin{matrix}\frac{36}{T_{s}^{2}} & 0\end{matrix}\right] \bmatrixx{r \\ y}
\end{align*}
with inputs $[r, y]$ and output $u$.

\section{Numerical Comparisons} \label{sec:numerical}
This section will demonstrate the equivalence between linear ADRC and the proposed equivalent PI(D) controllers numerically, using the same plant model as the one used in \cite{herbst2013simulative}. The plant model is a first-order system with $T = 1$ and $K = 1$, and the ADRC controller is tuned with $T_s = 1$ and $g = 10$. The parameters for the PI(D) controllers are found using \cref{eq:kp,eq:ki,eq:Tf}.

@@

\subsection{First-Order System}

@@

\subsection{Second-Order System}
The second-order system used in the experiment section in \cite{herbst2013simulative} is given by
% K = 1
% D = 1
% T = 1
% P = tf([K],[T^2, 2D*T, 1])
$$P(s) = \dfrac{1}{T^2s^2 + 2DTs + 1} = \dfrac{1}{s^2 + 2s + 1}$$

We show the same figures as for the first-order system
\begin{itemize}
	\item \cref{fig:second_order_K} shows the step response as the value of $K$ varies within the set $K \in \{0.1, 0.2, 0.5, 1, 2, 5\}$.
	\item \cref{fig:second_order_T} shows the step response as the value of $T$ varies within the set $T \in \{0.1, 0.2, 0.5, 1, 2, 5\}$.
	\item \cref{fig:second_order_bode_C} shows the Bode plot of the controllers.
	\item \cref{fig:second_order_bode_ry} shows the Bode plot of the closed-loop response from references to output.
	\item \cref{fig:second_order_bode_uy} shows the Bode plot of the closed-loop response from measurements to control signal.
\end{itemize}
The conclusions are largely the same for second-order systems. The response from measurements is identical between the ADRC controller and the equivalent PID controller. The response to references is slightly different, visible in the bode plot, but not different enough to be clearly visible in the step responses with varying plant parameters.

@@

\section{Second-order linear ADRC as a transfer function}
The transfer functions for the second-order ADRC controller (derived with symbolic computer calculations, reproduced in the software repository \cite{repo}) can be realized by the following PID+filter parameters
\begin{align}
	k_p &= \dfrac{(72g^3 + 108g^2)}{(3T_s^2 g^2 + 6T_s^2 g + T_s^2)} \\
	k_i &= \dfrac{216g^3}{(3T_s^3 g^2 + 6T_s^3 g + T_s^3)} \\
	k_d &= \dfrac{(6g^3 + 36g^2 + 18g)}{(3T_s g^2 + 6T_s g + T_s)} \\
	T_f &= T_s/6 \sqrt{1/(3g^2 + 6g + 1)} \\
	d &= (3g + 2)/2 \sqrt{1/(3g^2 + 6g + 1)} \\
	b &= (6/T_s)^2 / k_p
\end{align}
